---
title: "Testing DSM estimates of initial SOC"
author: "Francis Durnin-Vermette"
date: "2023/01/24"
output:
  html_document:
    toc: true
    number_sections: true
    df_print: paged
bibliography: bib.bib
editor_options:
  chunk_output_type: inline
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "doc") })
---

```{r include=FALSE}

library(ggplot2)
library(dplyr)
library(here)
library(stringr)

dir.create(tempdir()) #This fixes a bug if the temporary directory is not found

here::i_am("qc_initialc.rmd")
```

The purpose of this report is to visualize the realism of initial SOC estimates for Canadian long-term experimental sites.
These estimates were obtained using a digital soil mapping (DSM) method because some sites do not have an initial (meaning before the beginning of the experiment) SOC measurement, which is an important parameter in SOC modelling.

# Assumptions

1. In this report I will only look at sites which do not have measurements of initial SOC, because these are the only sites where DSM estimates of SOC are required.

<!-- 2. Many of these experiments did not sample SOC at depths 0-30cm. -->
<!-- We corrected these values using the following equation: -->
<!-- $$ -->
<!-- SOC_{30} = SOC_{measured} * \frac{30cm}{max_{depth} - min_{depth}} -->
<!-- $$ -->

2. Many of these experiments did not sample SOC at depths 0-30cm.
We corrected these values using the following algorithm:
a) For that treatment, sum the SOC depth measurements in order to achieve an estimate of the largest possible cross-section up to 30cm.
For instance, sum the measurements of 0-7.5cm and 7.5-15cm to calculate 0-15cm
b) Take this value and correct it up to 20cm
c) Take the control value from the same experiment, and correct the measurement for the 20-30cm depth and add this to the value.
d) This leaves us with a 0-30cm SOC estimate where the top 20cm change according to the treatment, and the lowest 10cm is unchanging / set to the control.

In order to apply depth corrections, we used the following equation:
$$
SOC_{depth_{target}} = SOC_{measured} * \frac{depth_{target}}{depth_{max} - depth_{min}}
$$

3. In order to visualize the change in SOC between the initial C estimate and the first measurement of SOC in the experiment, we need to decide on a year that the initial SOC estimate represents.
Here is my approach:
If the first SOC measurement in the experiment is after or during 2000, then I set the initial SOC to be the value in 1999.
If the first SOC measurement in the experiment is before 2000, then I assume that the initial C represents the SOC for the year *before* the first SOC measurement.

# Read data and join tables

Handle SOC measurement depth correction.

It turns out to be difficult to determine which treatment is a control based on the treatment name alone - especially for rotation trials.
We assume that a treatment is a control if the treatment_number is 1.
```{r, results='hide', message=FALSE, warning=FALSE}
site_data_all <- readr::read_csv(here("data/all_experiments_dummy.csv")) %>%
	mutate(soil_depth = paste0(soil_depth_min_cm,"-",soil_depth_max_cm))

bb <- site_data_all %>%
	select(location_name, field_name, year_name, treatment_number,treatment_name, replication_number, soil_depth,
				 soil_depth_min_cm, soil_depth_max_cm, soc_tha, soc_tha_30cm,
				 initial_soc_tha, initial_soc_year) %>%
	filter(!is.na(soc_tha)) %>%
	filter(!is.na(treatment_number)) %>%
	group_by(location_name, field_name, year_name, treatment_number, treatment_name, soil_depth_min_cm, soil_depth_max_cm,initial_soc_year) %>%
	
	
	summarise(across(soc_tha:initial_soc_tha, mean)) %>%
	group_by(year_name, field_name, soil_depth_min_cm, soil_depth_max_cm) %>%
	# Find bottom 10cm of soil profile SOC using control from that treatment
	mutate(soc_tha_control = filter(cur_data(),treatment_number==1)$soc_tha,
			 soc_tha_10cm_control = soc_tha_control*(10/(soil_depth_max_cm - soil_depth_min_cm))) %>%
			 ungroup(soil_depth_min_cm, soil_depth_max_cm) %>%
	group_by(location_name, treatment_name, .add=TRUE) %>%
	# Compute the total measured depth for each treatment, use this to estimate 20cm SOC
			 	summarise(lowest_depth = -sum(-soil_depth_max_cm, soil_depth_min_cm),
			 						soc_lowestdepth_dynamic = sum(soc_tha),
			 						soc_30cm_naive = mean(soc_tha_30cm),
			 						soc_10cm_control = mean(soc_tha_10cm_control)) %>%
			 	mutate(soc_20cm_dynamic = soc_lowestdepth_dynamic*(20/lowest_depth),
			 				 soc_20cm_naive = soc_30cm_naive*(20/30),
			 				 soc_30cm_dynamic_control = soc_20cm_dynamic + soc_10cm_control,
			 				 soc_30cm_naive_control = soc_20cm_naive + soc_10cm_control)
			 
```
			 
			 ```{r, results='hide', message=FALSE, warning=FALSE}

			 initialc_all <- readr::read_csv(here("data/lte_soc_30_cm_dsm.csv")) %>%
			 	rowwise() %>%
			 	mutate(location_name = stringr::str_split(location_n, ", ")[[1]][1])
			 
			 site_data_all_joined <- full_join(site_data_all, initialc_all, by="location_name") %>%
			 	# Add a column for soil depth (a factor in Lethbridge)
			 	mutate(soil_depth = paste0(soil_depth_min_cm,"-",soil_depth_max_cm)) %>%
			 	select(location_name,
			 				 initial_soc_tha, initial_dsm = `Can_SOC30 (Mg ha-1)`, initial_soc_year,
			 				 field_name, plot_id, block, treatment_number, treatment_name, replication_number, soil_depth,
			 				 year_name, soc_tha_30cm) %>%
			 	# Choose the measure of initial C we want to use
			 	mutate(initial_true = ifelse(is.na(initial_soc_tha), initial_dsm, initial_soc_tha), .before = initial_soc_tha) %>%
			 	group_by(location_name, field_name, plot_id, block, treatment_number, treatment_name, replication_number, soil_depth) %>%
			 	# Only look at sites where we don't already have measured initial C
			 	filter(is.na(initial_soc_year)) %>%
			 	# Set the initial C to either be 1 year before the first year in the experiment, or 1999 (maximum start date).
			 	mutate(initial_soc_year = ifelse(first(year_name) < 2000, first(year_name)-1, 1999)) %>%
			 	filter(!is.na(soc_tha_30cm))
			 site_data_all_joined

```
			 
			 # Graphing
			 
			 ## All sites which do not have measured initial C
			 
			 ```{r, message=FALSE, warning=FALSE, echo=FALSE}
			 site_data_all_joined %>%
			 	ggplot(aes(x=year_name, y=soc_tha_30cm)) +
			 	geom_point() +
			 	geom_point(aes(y=initial_true, x=initial_soc_year), color="red") +
			 	theme_bw() +
			 	facet_wrap(vars(location_name)) +
			 	ylab("SOC (Mg/ha)") +
			 	xlab("Year") +
			 	theme(aspect.ratio = 1)
			 ```
			 <!-- ## Ellerslie -->
			 <!-- ```{r} -->
			 <!-- site_data_all_joined %>% -->
			 <!-- 	filter(location_name == "Ellerslie") %>% -->
			 <!-- 	ggplot(aes(x=year_name, y=soc_tha_30cm)) + -->
			 <!-- 	geom_point() + -->
			 <!-- 	geom_point(aes(y=initial_true, x=initial_soc_year), color="red") + -->
			 <!-- 	theme_bw() + -->
			 <!-- 	facet_wrap(vars(treatment_name)) + -->
			 <!-- 	ylab("SOC (Mg/ha)") + -->
			 <!-- 	xlab("Year") + -->
			 <!-- 	ggtitle("Ellerslie") -->
			 <!-- ``` -->
			 <!-- ## Harrow -->
			 <!-- ```{r} -->
			 <!-- site_data_all_joined %>% -->
			 <!-- 	filter(location_name == "Harrow") %>% -->
			 <!-- 	ggplot(aes(x=year_name, y=soc_tha_30cm)) + -->
			 <!-- 	geom_point() + -->
			 <!-- 	geom_point(aes(y=initial_true, x=initial_soc_year), color="red") + -->
			 <!-- 	geom_segment(aes(y=initial_true, x=initial_soc_year, yend=200, xend=nth(year_name, n = 1))) + -->
			 <!-- 	theme_bw() + -->
			 <!-- 	facet_wrap(vars(treatment_name)) + -->
			 <!-- 	ylab("SOC (Mg/ha)") + -->
			 <!-- 	xlab("Year") + -->
			 <!-- 	ggtitle("Harrow") -->
			 <!-- ``` -->
			 
			 ## Lethbridge
			 
			 The DSM values for Lethbridge work well with the measured values.
			 The difference between the initial value and the first measured value never exceeds 2%/year.
			 
			 Lethbridge is an instance of a site where SOC was measured at depths 0-7.5, 0-15, 7.5-15, and 15-30cm.
			 In every case, SOC at 0-30cm was gap-filled using the above equation.
			 ```{r, message=FALSE, warning=FALSE, fig.width = 10, fig.height = 10, echo=FALSE}
			 lethbridge_increase <- site_data_all_joined %>%
			 	filter(location_name == "Lethbridge") %>%
			 	filter(!is.na(treatment_number)) %>%
			 	group_by_all() %>%
			 	ungroup(soil_depth, plot_id) %>%
			 	summarise_all(mean) %>%
			 	ungroup(replication_number) %>%
			 	summarise_all(mean) %>%
			 	group_by(treatment_name) %>%
			 	mutate(first_soc = first(soc_tha_30cm),
			 				 first_soc_year = first(year_name)) %>%
			 	select(treatment_name, first_soc, first_soc_year, initial_true, initial_soc_year) %>%
			 	unique %>%
			 	mutate(increase = first_soc - initial_true,
			 				 increase_per_year = increase / (first_soc_year - initial_soc_year),
			 				 fraction_increase = increase/initial_true,
			 				 fraction_increase_per_year = fraction_increase / (first_soc_year - initial_soc_year),
			 				 sign = ifelse(fraction_increase_per_year<0, "-", "+"))
			 
			 site_data_all_joined %>%
			 	filter(location_name == "Lethbridge") %>%
			 	filter(!is.na(treatment_number)) %>%
			 	# Get mean values between replicates and soil depth measurements. Lethbridge requires a particularly difficult grouping scheme
			 	group_by_all() %>%
			 	ungroup(soil_depth, plot_id) %>%
			 	summarise_all(mean) %>%
			 	ungroup(replication_number) %>%
			 	summarise_all(mean) %>%
			 	ggplot(aes(x=year_name, y=soc_tha_30cm)) +
			 	geom_segment(data=lethbridge_increase, aes(y=initial_true, x=initial_soc_year, yend=first_soc, xend=first_soc_year), color="red",
			 							 arrow = arrow(length = unit(0.1, "inches"))) +
			 	geom_text(data=lethbridge_increase, aes(x=first_soc_year, y=first_soc, hjust=ifelse(first_soc_year < 1990, -0.1, 1.1), label=paste0(sign,round(increase_per_year, 2)," Mg/ha/yr\n",sign,round(fraction_increase_per_year,2)*100, "%/yr")),
			 						color="red") +
			 	geom_point() +
			 	geom_point(aes(y=initial_true, x=initial_soc_year), color="red") +
			 	theme_bw() +
			 	facet_wrap(vars(treatment_name)) +
			 	ylab("SOC (Mg/ha)") +
			 	xlab("Year") +
			 	ggtitle("Lethbridge") +
			 	theme(aspect.ratio = 1)
			 ```
			 
			 ## Ottawa
			 
			 The DSM initial C estimate does not work as well for our Ottawa data - there are large 40-50% increases in SOC from the initial C to the first measured value.
			 I suspect this is because I am setting the initial C year to the year before the first measurement.
			 
			 These values would work better if we consider this to be the initial C ~10 years before the beginning of the experiment. This is shown in blue.
			 ```{r, message=FALSE, warning=FALSE, echo=FALSE}
			 ottawa_increase <- site_data_all_joined %>%
			 	filter(location_name == "Ottawa") %>%
			 	group_by(treatment_name) %>%
			 	mutate(first_soc = first(soc_tha_30cm),
			 				 first_soc_year = first(year_name)) %>%
			 	select(treatment_name, first_soc, first_soc_year, initial_true, initial_soc_year) %>%
			 	unique %>%
			 	mutate(increase = first_soc - initial_true,
			 				 increase_per_year = increase / (first_soc_year - initial_soc_year),
			 				 fraction_increase = increase/initial_true,
			 				 fraction_increase_per_year = fraction_increase / (first_soc_year - initial_soc_year),
			 				 sign = ifelse(fraction_increase_per_year<0, "-", "+"),
			 				 
			 				 recommended_initial_soc_year = initial_soc_year-10,
			 				 recommended_increase_per_year = increase / (first_soc_year - recommended_initial_soc_year),
			 				 recommended_fraction_increase_per_year = recommended_increase_per_year/initial_true)
			 
			 site_data_all_joined %>%
			 	filter(location_name == "Ottawa") %>%
			 	ggplot(aes(x=year_name, y=soc_tha_30cm)) +
			 	geom_segment(data=ottawa_increase, aes(y=initial_true, x=initial_soc_year, yend=first_soc, xend=first_soc_year), color="red",
			 							 arrow = arrow(length = unit(0.1, "inches"))) +
			 	geom_text(data=ottawa_increase, aes(x=first_soc_year, y=first_soc, label=paste0(sign,round(increase_per_year, 2)," Mg/ha/yr\n",sign,round(fraction_increase_per_year,2)*100, "%/yr")),
			 						color="red", hjust=-0.1) +
			 	geom_segment(data=ottawa_increase, aes(y=initial_true, x=recommended_initial_soc_year, yend=first_soc, xend=first_soc_year), color="blue",
			 							 arrow = arrow(length = unit(0.1, "inches"))) +
			 	geom_text(data=ottawa_increase, aes(x=recommended_initial_soc_year, y=initial_true, label=paste0(sign,round(recommended_increase_per_year, 2)," Mg/ha/yr\n",sign,round(recommended_fraction_increase_per_year,2)*100, "%/yr")),
			 						color="blue", hjust=-0.1, vjust=0) +
			 	geom_point() +
			 	geom_point(aes(y=initial_true, x=initial_soc_year), color="red") +
			 	theme_bw() +
			 	facet_wrap(vars(treatment_name)) +
			 	ylab("SOC (Mg/ha)") +
			 	xlab("Year") +
			 	ggtitle("Ottawa") +
			 	theme(aspect.ratio = 1)
			 ```
			 
			 ## Scott
			 
			 The DSM initial C values do not work very well for the Scott experiments.
			 These are very large increases between the initial C value and the first measurement of SOC.
			 
			 More reasonable increases (about +5%/year) are seen if we instead assume that the initial C represents the case 15 years before the first SOC measurement in the experiment (shown in blue).
			 
			 SOC measurements in Scott are taken at depths 0-7.5, 7.5-15, and 0-15cm.
			 We gap-filled these values to 0-30cm using the equation above.
			 ```{r, message=FALSE, warning=FALSE, fig.width = 10, fig.height = 10, echo=FALSE}
			 scott_increase <- site_data_all_joined %>%
			 	filter(location_name == "Scott") %>%
			 	group_by_all() %>%
			 	ungroup(soil_depth, plot_id) %>%
			 	summarise_all(mean) %>%
			 	ungroup(replication_number) %>%
			 	summarise_all(mean) %>%
			 	group_by(treatment_name) %>%
			 	mutate(first_soc = first(soc_tha_30cm),
			 				 first_soc_year = first(year_name)) %>%
			 	select(treatment_name, first_soc, first_soc_year, initial_true, initial_soc_year) %>%
			 	unique %>%
			 	mutate(increase = first_soc - initial_true,
			 				 increase_per_year = increase / (first_soc_year - initial_soc_year),
			 				 fraction_increase = increase/initial_true,
			 				 fraction_increase_per_year = fraction_increase / (first_soc_year - initial_soc_year),
			 				 sign = ifelse(fraction_increase_per_year<0, "-", "+"),
			 				 
			 				 recommended_initial_soc_year = initial_soc_year-15,
			 				 recommended_increase_per_year = increase / (first_soc_year - recommended_initial_soc_year),
			 				 recommended_fraction_increase_per_year = recommended_increase_per_year/initial_true)
			 
			 site_data_all_joined %>%
			 	filter(location_name == "Scott") %>%
			 	group_by_all() %>%
			 	ungroup(soil_depth, plot_id) %>%
			 	summarise_all(mean) %>%
			 	ungroup(replication_number) %>%
			 	summarise_all(mean) %>%
			 	ggplot(aes(x=year_name, y=soc_tha_30cm)) +
			 	geom_segment(data=scott_increase, aes(y=initial_true, x=initial_soc_year, yend=first_soc, xend=first_soc_year), color="red",
			 							 arrow = arrow(length = unit(0.1, "inches"))) +
			 	geom_text(data=scott_increase, aes(x=first_soc_year, y=first_soc, label=paste0(sign,round(increase_per_year, 2)," Mg/ha/yr\n",sign,round(fraction_increase_per_year,2)*100, "%/yr")),
			 						color="red", hjust=-0.1, size=3) +
			 	geom_segment(data=scott_increase, aes(y=initial_true, x=recommended_initial_soc_year, yend=first_soc, xend=first_soc_year), color="blue",
			 							 arrow = arrow(length = unit(0.1, "inches"))) +
			 	geom_text(data=scott_increase, aes(x=recommended_initial_soc_year, y=initial_true, label=paste0(sign,round(recommended_increase_per_year, 2)," Mg/ha/yr\n",sign,round(recommended_fraction_increase_per_year,2)*100, "%/yr")),
			 						color="blue", hjust=-0.1, vjust=0, size=3) +
			 	geom_point() +
			 	geom_point(aes(y=initial_true, x=initial_soc_year), color="red") +
			 	theme_bw() +
			 	facet_wrap(vars(treatment_name)) +
			 	ylab("SOC (Mg/ha)") +
			 	xlab("Year") +
			 	ggtitle("Scott") +
			 	theme(aspect.ratio = 1)
			 ```
			 
			 ## Swift Current New
			 
			 The DSM initial C values do not work very well for the Swift Current New experiments.
			 These are very large increases between the initial C value and the first measurement of SOC.
			 
			 More realistic values of % C increase (about 2-3% MgC/ha/yr) if we assume that the initial C value represents the SOC at 10 years before the beginning of the experiment (shown in blue).
			 ```{r, message=FALSE, warning=FALSE, fig.width = 10, fig.height = 10, echo=FALSE}
			 swiftcurrentnew_increase <- site_data_all_joined %>%
			 	filter(location_name == "Swift Current New") %>%
			 	group_by_all() %>%
			 	ungroup(field_name, soil_depth, plot_id) %>%
			 	summarise_all(mean) %>%
			 	ungroup(replication_number) %>%
			 	summarise_all(mean) %>%
			 	group_by(treatment_name) %>%
			 	mutate(first_soc = first(soc_tha_30cm),
			 				 first_soc_year = first(year_name)) %>%
			 	select(treatment_name, first_soc, first_soc_year, initial_true, initial_soc_year) %>%
			 	unique %>%
			 	mutate(increase = first_soc - initial_true,
			 				 increase_per_year = increase / (first_soc_year - initial_soc_year),
			 				 fraction_increase = increase/initial_true,
			 				 fraction_increase_per_year = fraction_increase / (first_soc_year - initial_soc_year),
			 				 sign = ifelse(fraction_increase_per_year<0, "-", "+"),
			 				 
			 				 recommended_initial_soc_year = initial_soc_year-10,
			 				 recommended_increase_per_year = increase / (first_soc_year - recommended_initial_soc_year),
			 				 recommended_fraction_increase_per_year = recommended_increase_per_year/initial_true) %>%
			 	summarise_all(first)
			 
			 site_data_all_joined %>%
			 	filter(location_name == "Swift Current New") %>%
			 	group_by_all() %>%
			 	ungroup(field_name, soil_depth, plot_id) %>%
			 	summarise_all(mean) %>%
			 	ungroup(replication_number) %>%
			 	summarise_all(mean) %>%
			 	ggplot(aes(x=year_name, y=soc_tha_30cm)) +
			 	geom_segment(data=swiftcurrentnew_increase, aes(y=initial_true, x=initial_soc_year, yend=first_soc, xend=first_soc_year), color="red",
			 							 arrow = arrow(length = unit(0.1, "inches"))) +
			 	geom_text(data=swiftcurrentnew_increase, aes(x=first_soc_year, y=first_soc, label=paste0(sign,round(increase_per_year, 2)," Mg/ha/yr\n",sign,round(fraction_increase_per_year,2)*100, "%/yr")),
			 						color="red", hjust=-0.1) +
			 	geom_segment(data=swiftcurrentnew_increase, aes(y=initial_true, x=recommended_initial_soc_year, yend=first_soc, xend=first_soc_year), color="blue",
			 							 arrow = arrow(length = unit(0.1, "inches"))) +
			 	geom_text(data=swiftcurrentnew_increase, aes(x=recommended_initial_soc_year, y=initial_true, label=paste0(sign,round(recommended_increase_per_year, 2)," Mg/ha/yr\n",sign,round(recommended_fraction_increase_per_year,2)*100, "%/yr")),
			 						color="blue", hjust=-0.1, vjust=0) +
			 	geom_point() +
			 	geom_point(data=swiftcurrentnew_increase,aes(y=initial_true, x=initial_soc_year), color="red") +
			 	theme_bw() +
			 	facet_wrap(vars(treatment_name)) +
			 	ylab("SOC (Mg/ha)") +
			 	xlab("Year") +
			 	ggtitle("Swift Current New") +
			 	theme(aspect.ratio = 1)
			 ```
			 
			 ## Swift Current Old
			 
			 The DSM initial C estimates generally work well for Swift Current Old.
			 Most treatments see increases of about +1-4% MgC/ha/yr SOC from the initial C estimate to the first year of SOC sampling.
			 
			 There are, however, some instances where this increase is in the range of 10-15%, which is less realistic.
			 These instances of large increases are the result of my assumption that the initial C is one year before the beginning of the experiment.
			 If we follow another assumption to decide the initial C year, then we might achieve more realistic changes in SOC at the beginning of the experiment.
			 
			 There are some instances where the % increase arrow in these graphs does not indicate the correct measured value.
			 I think this happens because Swift Current data does not have consistent entries for field_name.
			 
			 ```{r, message=FALSE, warning=FALSE, fig.width = 10, fig.height = 10, echo=FALSE}
			 swiftcurrentold_increase <- site_data_all_joined %>%
			 	filter(location_name == "Swift Current Old") %>%
			 	group_by_all() %>%
			 	ungroup(field_name, soil_depth) %>%
			 	summarise_all(mean) %>%
			 	ungroup(replication_number) %>%
			 	summarise_all(mean) %>%
			 	group_by(treatment_name) %>%
			 	mutate(first_soc = first(soc_tha_30cm),
			 				 first_soc_year = first(year_name)) %>%
			 	select(treatment_name, first_soc, first_soc_year, initial_true, initial_soc_year) %>%
			 	unique %>%
			 	mutate(increase = first_soc - initial_true,
			 				 increase_per_year = increase / (first_soc_year - initial_soc_year),
			 				 fraction_increase = increase/initial_true,
			 				 fraction_increase_per_year = fraction_increase / (first_soc_year - initial_soc_year),
			 				 sign = ifelse(fraction_increase_per_year<0, "-", "+")) %>%
			 	summarise_all(first)
			 
			 site_data_all_joined %>%
			 	filter(location_name == "Swift Current Old") %>%
			 	group_by_all() %>%
			 	ungroup(field_name, soil_depth) %>%
			 	summarise_all(mean) %>%
			 	ungroup(replication_number) %>%
			 	summarise_all(mean) %>%
			 	ggplot(aes(x=year_name, y=soc_tha_30cm)) +
			 	geom_segment(data=swiftcurrentold_increase, aes(y=initial_true, x=initial_soc_year, yend=first_soc, xend=first_soc_year), color="red",
			 							 arrow = arrow(length = unit(0.1, "inches"))) +
			 	geom_text(data=swiftcurrentold_increase, aes(hjust=ifelse(first_soc_year < 2000, -0.1, 1.1), x=first_soc_year, y=first_soc, label=paste0(sign,round(increase_per_year, 2)," Mg/ha/yr\n",sign,round(fraction_increase_per_year,2)*100, "%/yr")),
			 						color="red") +
			 	geom_point() +
			 	geom_point(data=swiftcurrentold_increase,aes(y=initial_true, x=initial_soc_year), color="red") +
			 	theme_bw() +
			 	facet_wrap(vars(treatment_name)) +
			 	ylab("SOC (Mg/ha)") +
			 	xlab("Year") +
			 	ggtitle("Swift Current Old") +
			 	theme(aspect.ratio = 1)
			 ```