---
title: "ICBM r~e~ Calculator Walkthrough & Tests"
author: "Francis Durnin-Vermette"
date: "2022/11/22"
output:
  html_notebook:
    toc: true
    number_sections: true
    df_print: paged
bibliography: re.bib
editor_options:
  chunk_output_type: inline
---

```{r include=FALSE}
library(ggplot2)
library(knitr)
library(dplyr)
library(geojson)
library(tibble)
library(purrr)
library(stringr)
library(here)
library(lubridate)
library(tidyr)

dir.create(tempdir()) #This fixes a bug if the temporary directory is not found

here::i_am("re_walkthrough.rmd")
source(here::here("re_script.r"))
source(here::here("re_functions.r"))
source(here::here("run_icbm_and_re.r"))
source(here::here("get_climate_data.r"))

```

# Quick Start - ICBM + r~e~ Calculator
In order to run ICBM simulations that include the calculated r~e~ parameter from climate data, you can use the function `RunICBMAndRe` from the script `run_icbm_and_re.r`. The definitions of every input are given in the header of that script. The function returns the same output as the ICBM function `icbm_holos4_classic_manure`; a table with ICBM C pools for every year, as well as the yearly mean r~e~ value.

## Importing climate data
Climate data can be manually inputted into the model runs, or the function `getClimateData` can be used to automatically build the climate data input from a folder containing .csv files for each year of daily climate data.

Note that the length of the simulation run in years is determined by the number of years represented in the climate data.
The climate data should have these mandatory columns: `Year`, `JulianDay`, `Tavg`, which is the mean daily air temperature, `PREC`, which is the mean daily precipitation, and `PET`, which is the mean daily PET.

Note: use the `suppressWarnings` function to run a function without any warnings
(e.g. `suppressWarnings(run_icbm_and_re.r())`).

### using `getClimateData()`

```{r}
# Import climate data
get_climate_data_example <- getClimateData(
	climate_data_folder = "climate_data",
	master_data_folder = "master",
	site_name = "244014")
```


### Using manual input

```{r}
# Import weather data
# This is arbitrary weather data from the lat/long 45/45.
holos_nasa_climate_example <- readr::read_csv(here::here("holos_nasa_climate_example.csv")) %>%
	filter(Year >= 2000) %>%
	rename(Tavg = MeanDailyAirTemperature,
					 PREC = MeanDailyPrecipitation,
					 PET = MeanDailyPET)
```

## Running ICBM

```{r}
# These are arbitrary annual C inputs
iag <- rep(977,length(unique(get_climate_data_example$Year))) #aboveground young
ibg <- rep(288,length(unique(get_climate_data_example$Year))) #belowground young
iman <- rep(88,length(unique(get_climate_data_example$Year))) #manure
# This is annual yield for an arbitrary crop type
yield <- 3000
perennial <- FALSE
# These are arbitrary soil parameters;
# not related to an SLC polygon or the location of the climate data
SoilOrganicC_Percent <- 5
ClayContent <- 0.05
SandContent <- 0.2

RunICBMAndRe(yield = yield,
						 iag = iag,
						 ibg = ibg,
						 iman = iman,
						 ag_init = 0,
						 bg_init = 0,
						 o_init = 0,
						 DailyClimateTable = get_climate_data_example,
						 perennial = perennial,
						 SoilOrganicC_Percent = SoilOrganicC_Percent,
						 ClayContent = ClayContent,
						 SandContent = SandContent,
						 irrigation_use_estimate = FALSE)
```


# Quick Start - r~e~ Calculator
This code chunk represents the necessary input table and function call to use the r~e~ calculator.
Definitions of the input variables can be found in the header section of the re_script.r file.
`calculate_re` returns a single numeric value: the mean daily r~e~ value averages across every day of one year.
In order to build a vector of r~e~ values for a multi-year simulation, see the code chunks in [this section](#multiyear).

```{r}
quickstart_input <- tibble(
	Year = 2000,
	JulianDay = 1:365,
	Tavg = rep(0,365),
	PREC = rep(0,365),
	PET = rep(0,365)
)

calculate_re(quickstart_input,
						 yield=3000,
						 perennial=F,
						 SoilOrganicC_Percent = 5,
						 ClayContent = 0.05,
						 SandContent = 0.2,
						 SoilTopThickness = 250,
						 Temp_min = -3.78,
						 Temp_max = 30,
						 r_s = 0.42,
						 r_wp = 0.18,
						 ReferenceAdjustment = 0.10516,
						 r_c = NA,
						 tillage_soil = "Brown",
						 tillage_type = "Intensive Tillage",
						 irrigation_region = "Canada",
						 irrigation_use_estimate = TRUE,
						 irrigation = 0)
```


We discuss irrigation calculations [here](#irrigationdisc).

We discuss tillage calculations [here](#tillage).

Here is the full standalone script:
```{r}
calculate_re
```

# Walkthrough
## Input data

```{r, message=FALSE}
holos_re_calculation <- readr::read_csv(here::here("holos_re_calculation.csv"))

# climate data
daily_input_test <- tibble(
	Year = 2000,
	JulianDay = 1:365,
	Tavg = holos_re_calculation$InputTemperature,
	PREC = holos_re_calculation$InputPrecipitation,
	PET = holos_re_calculation$InputEvapotranspiration)

# other values
yield = 3000
perennial = FALSE
SoilOrganicC_Percent = 2
ClayContent = 0.05
SandContent = 0.2
alfa = 0.7
SoilTopThickness = 250
Temp_min = -3.78
Temp_max = 30
r_s = 0.42
r_wp = 0.18
ReferenceAdjustment = 0.10516
r_c = NA
tillage_soil = "Brown"
tillage_type = "Intensive Tillage"
irrigation_region = "Canada"
irrigation_use_estimate = FALSE
irrigation = 0
```



## Green area index (GAI) dynamics

### Eq. 2.2.1-1 through Eq. 2.2.1-3
$$GAI_{max}=0.0731\times(\frac{Yield}{1000})^{2}+0.408\times(\frac{Yield}{1000})$$
$$MidSeason=EmergenceDay+(\frac{RipeningDay-EmergenceDay}{2})$$
$$GAI=GAI_{max}\times e^{-[\frac{(JulianDay-MidSeason)^{2}}{2*Variance}]}$$

```{r}
calculateGAI
```

Where
`Yield` is the crop yield (kg dry matter ha-1),
`JulianDay` is the day,
`MidSeason` is the median day of growing season (Julian day),
`EmergenceDay` is the day of crop emergence (Julian day),
`RipeningDay` is the day of crop ripening (Julian day),
`GAI_max` is the maximum amplitude of green area index,
`GAI` is the green area index, and
`Variance` is the width of the distribution function

```{r}
GAI_return <- calculateGAI(daily_input_test$JulianDay,
													 yield,
													 perennial)
GAI_return
GAI <- GAI_return[["GAI"]]
```



## Water content at wilting point and field capacity

### Eq. 2.2.1-4 through Eq. 2.2.1-10
$$OrgC_{factor}=-0.837531+0.430183\times SoilOrganicC_{Percent}$$

$$Clay_{factor}=-1.40744+0.0661969\times ClayContent\times100$$

$$Sand_{factor}=-1.51866+0.0393284\times SandContent\times100$$

$$WiltingPoint_{Percent}=14.2568+7.36318×0.06865+0.108713×OrgC_{factor}-0.0157225×OrgC_{factor}2+0.00102805×OrgC_{factor}3+0.886569×Clay_{factor}-0.223581×OrgC_{factor}×Clay_{factor}+0.0126379×OrgC_{factor}2×Clay_{factor}-0.017059×Clay_{factor}2+0.0135266×OrgC_{factor}Clay_{factor}2-0.0334434×Clay_{factor}3-0.0535182×Sand_{factor}-0.0354271×OrgC_{factor}×Sand_{factor}-0.00261313×OrgC_{factor}2×Sand_{factor}-0.154563×Clay_{factor}×Sand_{factor}-0.0160219×OrgC_{factor}×Clay_{factor}×Sand_{factor}-0.0400606×Clay_{factor}2×Sand_{factor}-0.104875×Sand_{factor}2+0.0159857×OrgC_{factor}Sand_{factor}2-0.0671656×Clay_{factor}Sand_{factor}2-0.0260699×Sand_{factor}3$$

$$WiltingPoint=\frac{WiltingPoint_{Percent}}{100}$$
$$FieldCapacity_{Percent}=29.7528+10.3544×0.0461615+0.290955×OrgC_{factor}-0.0496845×OrgC_{factor}2+0.00704802×OrgC_{factor}3+0.269101×Clay_{factor}-0.176528×OrgC_{factor}×Clay_{factor}+0.0543138×OrgC_{factor}2×Clay_{factor}+0.1982×Clay_{factor}2-0.060699×Clay_{factor}3-0.320249×Sand_{factor}-0.0111693×OrgC_{factor}2×Sand_{factor}+0.14104×Clay_{factor}×Sand_{factor}+0.0657345×OrgC_{factor}×Clay_{factor}×Sand_{factor}-0.102026×Clay_{factor}2×Sand_{factor}-0.04012×Sand_{factor}2+0.160838×OrgC_{factor}Sand_{factor}2-0.121392×Clay_{factor}Sand_{factor}2-0.061667×Sand_{factor}3$$
$$FieldCapacity=\frac{FieldCapacity_{Percent}}{100}$$

```{r}
calculateWaterContent
```

Where
`OrgCfactor` is the organic carbon factor,
`SoilOrganicCPercent` is the percentage of organic carbon in soil, by weight (%, SLC data table),
`Clayfactor` is the clay factor,
`ClayContent` is the proportion of clay in soil (SLC data table),
`Sandfactor` is the sand factor,
`SandContent` is the proportion of sand in soil (SLC data table),
`WiltingPointPercent` is the water content at wilting point (%),
`WiltingPoint` is the proportion of water content at wilting point (mm^3^ mm^-3^),
`FieldCapacityPercent` is the water content at field capacity (%), and
`FieldCapacity` is the proportion of water at field capacity (mm^3^ mm^-3^).

```{r}
WaterContent_return <- calculateWaterContent(SoilOrganicC_Percent = SoilOrganicC_Percent,
																						 ClayContent = ClayContent,
																						 SandContent = SandContent)
WaterContent_return
WiltingPoint <- WaterContent_return[["WiltingPoint"]]
FieldCapacity <- WaterContent_return[["FieldCapacity"]]
```


## Soil temperature

### Eq. 2.2.1-11 & Eq. 2.2.1-12

$$SoilMeanDepth=\frac{SoilTopThickness}{20}$$
$$LeafAreaIndex=0.8\times GAI$$

```{r}
SoilTopThickness <- 250
SoilMeanDepth <- SoilTopThickness/20
LeafAreaIndex <- 0.8*GAI
```

Where
`SoilMeanDepth` is the Soil mean depth (mm),
`SoilTopThickness` is the thickness of top layer (mm), and
`LeafAreaIndex` is the leaf area index

## Surface temperature
```{r}
Temperature <- daily_input_test$Tavg
```

###  Eq. 2.2.1-13 & Eq. 2.2.1-14
If Temperature < 0$^\circ$C
$$SurfaceTemp=0.20\times Temperature$$
If Temperature >= 0$^\circ$C
$$SurfaceTemp=Temperature\times(0.95+0.05\times e^{-0.4\times(LeafAreaIndex-3)})$$

```{r}
SurfaceTemp <- ifelse(Temperature < 0, 0.20*Temperature,
											Temperature*(0.95+0.05*exp(-0.4*(LeafAreaIndex-3))))
```

Where
`Temperature` is the daily mean air temperature (˚C, by month - SLC data table), and
`SurfaceTemp` is the soil surface temperature (˚C)

## Soil Temperature

### Sidenote: handling recursive equations
How are we able to recursively calculate soil temperature using the previous day's calculated soil temperature?
Our implementation is based on Dan Chaltiel's answer to this Stack Overflow question: https://stackoverflow.com/questions/67525729/recursive-lag-during-mutate

Inside the `mutate()` call, we are able to use the context-dependent expression `cur_data_all()` to cache the current iteration of the data.
This allows us to get the current row value for the column time using `db$time[row]`, and then multiply this by the previous value of the column we are accumulating.

We use this approach because it lets us keep the functions vectorized.

```{r}
sidenote1 <- tibble(time = 1:5) %>%
	mutate(
		c = accumulate(.x=time, .f=function(acc, row) {
			db = cur_data_all()
			time = db$time[row]
			acc*time
			}) %>% unlist()
	)
sidenote1
```


### Eq. 2.2.1-15 & Eq. 2.2.1-16

If JulianDay = 1
$$SoilTemp_{d}=0$$
If JulianDay>1
$$SoilTemp_{d}=SoilTemp_{d-1}+(SurfaceTemp-SoilTemp_{d-1}\times 0.24\times e^{-SoilMeanDepth\times 0.017}\times e^{-0.15\times GAI})$$

Using the same method as above, we can `accumulate()` the value of `SoilTemp` using its value in the previous row (`SoilTemp_dprev`).
In this case, the name "accumulate" might lead to some confusion because the soil temperature goes both up and down over the course of the year.
The function `accumulate()` doesn't itself increase the value of of a variable, it just passes both the previous value and the current value of the variable into a function that you specify.

The function we are using in `accumulate()` here is `SoilTemp_d = SoilTemp_dprev + (SurfaceTemp-SoilTemp_dprev) * 0.24 * exp(-SoilMeanDepth*0.017) * exp(-0.15*GAI)`

```{r}
calculateSoilTemp
```
Where
`SoilTemp` is the soil temperature (˚C), and
`SoilTemp_dprev` is the soil temperature on the previous day (˚C)

```{r}
SoilTemp <- calculateSoilTemp(SurfaceTemp = SurfaceTemp,
															GAI = GAI,
															SoilMeanDepth = SoilMeanDepth)
```

## Irrigation
Irrigation can either be calculated using Eqs. 2.2.1-17 and 2.2.1-18, or specified by the user.

### Monthly distribution of irrigation

From the Holos algorithm document:
"The proportion of the annual irrigation amount applied in each month and for each province is estimated as an average of values derived from the Agricultural Water Use Survey for 2010 and 2012 [@AgriculturalWaterUse2011; @AgriculturalWaterUse2013]".

```{r irrigationtable}
Irrigation_percentage_monthly <- tibble(
	`Province/Region` = c("Canada", "BC", "AB", "SK", "MB", "ON", "QC", "Atlantic Provinces"),
	`Jan` = 0,
	`Feb` = 0,
	`Mar` = 0,
	`Apr` = c(4.69, 8.00, 3.06, 15.08, 0.22, 3.81, 10.54, 7.30),
	`May` = c(4.69, 8.00, 3.06, 15.08, 0.22, 3.81, 10.54, 7.30),
	`Jun` = c(15.64, 19.54, 13.92, 14.80, 11.30, 18.35, 19.54, 10.62),
	`Jul` = c(38.96, 27.77, 42.65, 29.58, 48.15, 41.46, 32.24, 27.97),
	`Aug` = c(26.92, 25.66, 28.04, 19.40, 34.88, 26.74, 25.97, 34.06),
	`Sep` = c(4.56, 5.51, 4.63, 3.03, 2.61, 2.91, 2.71, 5.63),
	`Oct` = c(4.56, 5.51, 4.63, 3.03, 2.61, 2.91, 2.71, 5.63),
	`Nov` = 0,
	`Dec` = 0
) %>%
	mutate(Total = select(., where(is.numeric)) %>% rowSums())
Irrigation_percentage_monthly %>%
	kable(caption = "Percentage of total annual irrigation water applied by month for each province/region in Canada using average values across 2010 and 2012 (Adapted from the Holos algorithm document, Table 47).")
```


The order of the equations has been changed from the original Holos algorithm document because irrigation is needed to calculate water balance.

### Eq. 2.2.1-17 & Eq. 2.2.1-18

Irrigation is calculated using the built-in calculator if PE > P
$$Irrigation_{annual}=PE-P$$
$$Daily_{PrecipTotal}=P_{Daily}+(\frac{Irrigation_{annual}\times Fraction_{monthly}}{Days_{month}})$$

```{r}
P <- sum(daily_input_test$PREC)
PE <- sum(daily_input_test$PET)
Region <- "QC"

Days_month <- lubridate::days_in_month(as.Date(paste(daily_input_test$Year, daily_input_test$JulianDay, sep="-"), "%Y-%j"))

Fraction_monthly <- Irrigation_percentage_monthly %>%
	filter(`Province/Region` == Region) %>%
	`[`(names(Days_month)) %>%
	unlist %>%
	`/`(100)

if (PE > P) {
	warning("Using Eqs. 2.2.1-17 and 2.2.1-18 to calculate default irrigation values")
	Irrigation_annual <- PE - P
	
	Irrigation <- ((Irrigation_annual * Fraction_monthly)/Days_month) %>%
		unname
} else {
	Irrigation <- 0
}

```
Where
`PE` is the annual evapotranspiration (mm yr-1),
`P` is the annual precipitation (mm yr-1),
`Irrigation_annual` is the annual irrigation (mm yr-1),
`Fraction_monthly` is the assumed monthly distribution of irrigation (%) which is detailed in the next section,
`Days_month` is the number of days per month, and
`Irrigation` is the irrigation moisture input per day (mm d-1)

However, because Holos does not use the `if (PE > P)` conditional to assess whether irrigation should be added, we get a different result from the Holos version if we include irrigation.
For the sake of this test, we remove the influence of irrigation.
```{r}
Irrigation <- 0
```



## Crop Evapotranspiration
### Eq. 2.2.1-19 & Eq. 2.2.1-20

$$K_{c}=1.3-(1.3-0.8)\times e^{-0.17\times GAI}$$
$$ET_{c}=ET_{0}\times K_{c}$$

```{r}
#Eq. 2.2.1-19
K_c <- 1.3 - (1.3-0.8) * exp(-0.17*GAI)
ET_0 <- daily_input_test$PET
#Eq. 2.2.1-20
ET_c <- ET_0 * K_c
```

Where
`K_c` is the crop coefficient [@martelParsimoniousWaterBudget2021],
`ET_c` is the crop evapotranspiration under standard conditions (mm day^-1^), and
`ET_0` is the daily reference crop evapotranspiration (mm day^-1^), which should be calculated according to the FAO Penman-Monteith equation [@allenCropEvapotranspirationGuidelinesComputing1998]


## Soil Available Water

### Eq. 2.2.1-21 through Eq. 2.2.1-24

If Precipitation + Irrigation < 0.2*GAI
$$CropInterception=Daily_{PrecipTotal}$$
If Precipitation + Irrigation >= 0.2*GAI
$$CropInterception=0.2\times GAI$$
If CropInterception > ET~c~
$$CropInterception=ET_{c}$$

$$SoilAvailWater=Daily_{PrecipTotal}-CropInterception$$
```{r}
Precipitation <- daily_input_test$PREC

#Eq. 2.2.1-21 - Eq. 2.2.1-22
CropInterception <- ifelse(Precipitation + Irrigation < 0.2*GAI,
													Precipitation + Irrigation,
													0.2*GAI)

#Eq. 2.2.1-23
CropInterception <- ifelse(CropInterception > ET_c, ET_c, CropInterception)

#Eq. 2.2.1-24
SoilAvailWater <- Precipitation + Irrigation - CropInterception
```
Where
`CropInterception` is the crop interception of precipitation + irrigation (mm day^-1^),
`Precipitation` is the daily precipitation (mm day^-1^), and
`SoilAvailWater` is the soil available water (mm day^-1^)

## Water Balance

### Eq. 2.2.1-25 through Eq. 2.2.1-35

$$VolSoilWaterContent=\frac{WaterStorage_{d-1}}{SoilTopThickness}$$
If VolSoilWaterContent = 0 or NA
$$VolSoilWaterContent=WiltingPoint$$
$$K_{r}=(1-\frac{0.95\times FieldCapacity-VolSoilWaterContent}{0.95\times FieldCapacity-alfa\times WiltingPoint})^{2}$$
$$K_{r}=max(min(1,K_{r}),0)$$
If VolSoilWaterContent < alfa/100*WiltingPoint
$$K_{r}=0$$
$$ET_{a}=ET_{c}\times K_{r}$$
If JulianDay = 1
$$WaterStorage_{d-1}=FieldCapacity\times SoilTopThickness$$
$$WaterStorage_{d}=FieldCapacity\times SoilTopThickness$$
If JulianDay > 1
$$DeepPerc=WaterStorage_{d-1}-FieldCapacity\times SoilTopThickness$$
If DeepPerc < 0
$$DeepPerc = 0$$
$$WaterStorage_{d}=WaterSotrage_{d-1}+SoilAvailWater-ET_{a}-DeepPerc$$

We are interested in calculating topsoil water storage where the previous day's water storage is an input variable.
We use `mutate()` and `accumulate()` to vectorize water balance calculations.

In order to calculate water storage, we use this function:
```{r}
calculateWaterStorage
```
Where
`VolSoilWaterContent` is the volumetric soil water content  (mm^3^ mm^-3^),
`K_r` is the soil coefficient (dimensionless),
`alfa` is the minimum water storage fraction of WiltingPoint, and
`ET_a` is the actual evapotranspiration (mm day^-1^)
`WaterStorage_dprev` is the previous day's water storage in topsoil (mm day^-1^),
`WaterStorage_d` is the current day's water storage in topsoil (mm day^-1^), and
`DeepPerc` is the water lost to percolation down the soil profile (mm day^-1^)

```{r}
WaterStorage <- calculateWaterStorage(SoilAvailWater = SoilAvailWater,
																			ET_c = ET_c,
																			alfa = 0.7,
																			SoilTopThickness = SoilTopThickness,
																			WiltingPoint = WiltingPoint,
																			FieldCapacity = FieldCapacity)
```

## Decomposition rate - effect of soil temperature

Here, we assume that `SoilTemp_dprev` and `SoilTemp_d` are both equal to 0 on `JulianDay = 1`.

### Eq. 2.2.1-36 & Eq. 2.2.1-37
$$r_{e\_temp}=\frac{(SoilTemp_{d-1}-Temp_{min})^{2}}{(Temp_{max}-Temp_{min})^{2}}$$
If SoilTemp~d-1~ < -3.78$^\circ$C
$$r_{e\_temp}=0$$

```{r}
SoilTemp_dprev <- lag(SoilTemp, default=0)
Temp_min <- -3.78
Temp_max <- 30

# Eq. 2.2.1-36 - Eq. 2.2.1-37
re_temp <- ifelse(SoilTemp_dprev < -3.78, 0, ((SoilTemp_dprev - Temp_min)^2)/((Temp_max-Temp_min)^2))
```
Where
`re_temp` is the temperature response factor,
`Temp_min` is the critical soil temperature (˚C), and
`Temp_max` is the maximum soil temperature (˚C)

## Decomposition rate - effect of soil moisture
### Eq. 2.2.1-38 through Eq. 2.2.1-43

$$VolSoilWaterContent_{sat}=1.2\times FieldCapacity$$
$$VolSoilWaterContent_{opt}=0.9\times FieldCapacity$$
If VolSoilWaterContent > VolSoilWaterContent~opt~
$$r_{e\_water}=(1-r_{s})\times \frac{VolSoilWaterContent-VolSoilWaterContent_{opt}}{VolSoilWaterContent_{opt}-VolSoilWaterContent_{sat}}+1$$
If VolSoilWaterContent < WiltingPoint
$$r_{e\_water}=r_{wp}\times \frac{VolSoilWaterContent}{WiltingPoint}$$
If WiltingPoint <= VolSoilWaterContent <= VolSoilWaterContent~opt~
$$r_{e\_water}=(1-r_{wp})\times \frac{VolSoilWaterContent - WiltingPoint}{VolSoilWaterContent_{opt}-WiltingPoint}+r_{wp}$$
$$r_{e\_water}=max(min(1,r_{e\_water}),0)$$

```{r}
WaterStorage_dprev <- lag(WaterStorage, default = FieldCapacity[1]*SoilTopThickness)
VolSoilWaterContent <- calculateVolSoilWaterContent(WaterStorage_dprev, SoilTopThickness, WiltingPoint)

r_s <- 0.42
r_wp <- 0.18

# Eq. 2.2.1-38
VolSoilWaterContent_sat <- 1.2*FieldCapacity
# Eq. 2.2.1-39
VolSoilWaterContent_opt <- 0.9*FieldCapacity

# Eq. 2.2.1-40 - Eq. 2.2.1-42
re_water <- ifelse(
	# if...
	VolSoilWaterContent > VolSoilWaterContent_opt,
	# then...
	(1 - r_s)*((VolSoilWaterContent-VolSoilWaterContent_opt)/(VolSoilWaterContent_opt-VolSoilWaterContent_sat)) + 1, # Eq. 2.2.1-40
	ifelse(
		# elseif...
		VolSoilWaterContent < WiltingPoint,
		# then...
		r_wp*((VolSoilWaterContent)/(WiltingPoint)), # Eq. 2.2.1-41
		# else...
		(1-r_wp)*((VolSoilWaterContent-WiltingPoint)/(VolSoilWaterContent_opt-WiltingPoint)) + r_wp)) # Eq. 2.2.1-42

# Eq. 2.2.1-43
re_water <- pmin(pmax(0,re_water),1)
```
Where
`VolSoilWaterContent_sat` is the volumetric soil water content at saturation level (mm^3^ mm^-3^),
`VolSoilWaterContent_opt` is the volumetric soil water content at optimal water content (mm^3^ mm^-3^),
`r_s` is the reference saturation (mm^3^ mm^-3^),
`r_wp` is the reference wilting point (mm^3^ mm^-3^), and
`re_water` is the moisture response factor

## Climate Factor (re~crop~)

### Eq. 2.2.1-44 through Eq. 2.2.1-46

$$r_{e\_x1}=r_{e\_temp}\times r_{e\_water}$$
$$r_{e\_crop\_daily}=\frac{r_{e\_x1}}{ReferenceAdjustment}$$
$$r_{e\_crop}=mean(r_{e\_crop\_daily})$$

In Holos, re~crop~ is calculated daily and then averaged annually.
```{r}
ReferenceAdjustment = 0.10516

# Eq. 2.2.1-44
re_x1 <- re_temp*re_water

# Eq. 2.2.1-45
re_cropdaily <- re_x1/ReferenceAdjustment

# Eq. 2.2.1-46
re_crop <- mean(re_cropdaily)
```
Where
`re_x1` is the uncalibrated daily climate parameter,
`ReferenceAdjustment` is the calibration factor for a bare-fallow treatment considering a soil thickness of 25 cm (Uppsala, Sweden),
`re_cropdaily` is the daily climate parameter, and
`re_crop` is the climate parameter


## Tillage Factor (r~c~)
From the Holos algorithm document: "For the prairie provinces (Alberta, Saskatchewan, and Manitoba), changes in tillage management have been shown to influence soil carbon stocks. Accordingly, the tillage factor varies with soil type and tillage intensity."

For non-prairie provinces, `r_c = 1.0`,
for perennials, `r_c = 0.9` for the first year of planting; then the no-till factor is used for subsequent years,
for root crops in every province and soil zone, `r_c = 1.13`.
```{r}
tibble(`Soil type` = c("Brown", "Dark Brown", "Black"),
			 `Intensive Tillage` = c(1,1,1),
			 `Reduced tillage` = c(0.9, 0.85, 0.8),
			 `No-till` = c(0.8, 0.7, 0.6)) %>%
	kable(caption="r_c factor for Alberta, Saskatchewan, and Manitoba only (adapted from Holos algorithm document)")
```

```{r}
r_c <- 1
```


## Climate/management Factor (r~e~)

### Eq. 2.2.1-47

$$r_{e}=r_{c}\times r_{e\_crop}$$

```{r}
r_edaily <- re_cropdaily*r_c
r_e <- mean(r_edaily)
r_e
```
Where
`r_c` is the tillage factor, and
`re_cropdaily` is the daily climate/management factor,
`r_edaily` is the daily climate/management factor, and
`r_e` is the climate/management factor averaged for the year


# Testing

## This notebook vs. Holos
```{r, include = FALSE}
daily_input_test_all <- daily_input_test %>%
	bind_cols(Days_month = Days_month,
						Fraction_monthly = Fraction_monthly,
						Irrigation = Irrigation,
						
						GAI = GAI,
						
						SurfaceTemp = SurfaceTemp,
						SoilTemp = SoilTemp,
						
						K_c = K_c,
						ET_c = ET_c,
						Days_month = Days_month,
						Fraction_monthly = Fraction_monthly,
						Irrigation = Irrigation,
						CropInterception = CropInterception,
						SoilAvailWater = SoilAvailWater,
						WaterStorage = WaterStorage,
						SoilTemp_dprev = SoilTemp_dprev,
						WaterStorage_dprev = WaterStorage_dprev,
						VolSoilWaterContent = VolSoilWaterContent,
						
						
						re_temp = re_temp,
						re_water = re_water,
						re_x1 = re_x1,
						re_cropdaily = re_cropdaily,
						
						re_crop = re_crop,
						r_c = r_c,
						
						r_edaily = r_edaily,
						r_e = r_e)
```

r~e~ value from this notebook:
```{r}
test_re <- daily_input_test_all %>%
	summarise(r_e = mean(r_e)) %>%
	unlist
```

r~e~ value from Holos:
```{r}
holos_re <- holos_re_calculation %>%
	mutate(r_e = mean(ClimateParameter)) %>%
	summarise(r_e = mean(ClimateParameter)) %>%
	unlist
```

Percentage difference between the two implementations:
```{r}
PercentageDifference = 100*(holos_re - test_re)/holos_re

PercentageDifference
```
This means our R implementation of r~e~ is only `r PercentageDifference`% different from the r~e~ as calculated by Holos.

## re_script.r vs. Holos

Now let's test whether our standalone r~e~ script works as well:

```{r, include = FALSE}
source(here::here("re_script.r"))
```

```{r}
script_re <- calculate_re(daily_input_test,
													yield=3000,
													perennial=F,
													SoilOrganicC_Percent=2,
													ClayContent=0.05,
													SandContent=0.2,
													irrigation_use_estimate = FALSE)

PercentageDifference = 100*(holos_re - script_re)/holos_re

PercentageDifference
```
The error value from `script_re` (the r~e~ calculator implementation in the standalone script) is the same as from `test_re` (the r~e~ calculator implementation in this R notebook), meaning our re_script.r implementation works as intended.

## Testing irrigation parameters {#irrigationdisc}

In order to account for irrigation, we must specify whether we want to estimate daily irrigation values using Eqs. 2.2.1-45 and 2.2.1-46, or whether we want to put in our own daily irrigation values.

In order to use the built-in daily irrigation estimations, use `irrigation_use_estimate = TRUE` and `irrigation_region`:
```{r}
calculate_re(daily_input_test,
						 yield = 3000,
						 perennial = FALSE,
						 SoilOrganicC_Percent = 2,
						 ClayContent = 0.05,
						 SandContent = 0.2,
						 
						 irrigation_use_estimate = TRUE,
						 irrigation_region = "Canada")
```

If you do *not* want to use the built-in daily irrigation estimations, a vector of daily irrigation values for the year can be passed into the `irrigation` parameter directly with `irrigation_use_estimate` set to `FALSE`:

```{r}
calculate_re(daily_input_test,
						 yield = 3000,
						 perennial = FALSE,
						 SoilOrganicC_Percent = 2,
						 ClayContent = 0.05,
						 SandContent = 0.2,
						 
						 irrigation_use_estimate = FALSE,
						 irrigation = rep(20,365))
```

`irrigation_use_estimate` can only take values of either `TRUE` or `FALSE`. If it is not, an error is returned:

```{r}
try(calculate_re(daily_input_test,
								 yield = 3000,
								 perennial = FALSE,
								 SoilOrganicC_Percent = 2,
								 ClayContent = 0.05,
								 SandContent = 0.2,
						 
								 irrigation_use_estimate = NA))
```

An error is returned if `irrigation_region` is not passed a valid argument:

```{r}
try(calculate_re(daily_input_test,
								 yield = 3000,
								 perennial = FALSE,
								 SoilOrganicC_Percent = 2,
								 ClayContent = 0.05,
								 SandContent = 0.2,
								 
								 irrigation_region = "Nebraska",
								 irrigation_use_estimate = TRUE))
```

## Testing tillage parameter (r~c~) {#tillage}

By default, r~c~ is given a value of 1.
This can either be changed based on the tillage table by specifying a `tillage_type` and a `tillage_soil`, or by supplying a vector of `r_c` values of length 365.
```{r}
calculate_re(daily_input_test,
						 yield = 3000,
						 perennial = FALSE,
						 SoilOrganicC_Percent = 2,
						 ClayContent = 0.05,
						 SandContent = 0.2,
						 
						 tillage_type = "Reduced Tillage",
						 tillage_soil = "Brown")

calculate_re(daily_input_test,
						 yield = 3000,
						 perennial = FALSE,
						 SoilOrganicC_Percent = 2,
						 ClayContent = 0.05,
						 SandContent = 0.2,
						 
						 tillage_type = "No-till",
						 tillage_soil = "Brown")

calculate_re(daily_input_test,
						 yield = 3000,
						 perennial = FALSE,
						 SoilOrganicC_Percent = 2,
						 ClayContent = 0.05,
						 SandContent = 0.2,
						 
						 tillage_type = "No-till",
						 tillage_soil = "Black")

calculate_re(daily_input_test,
						 yield = 3000,
						 perennial = FALSE,
						 SoilOrganicC_Percent = 2,
						 ClayContent = 0.05,
						 SandContent = 0.2,
						 
						 r_c = rep(0.05,365))
```
An error is returned if either `tillage_type` or `tillage_soil` are not passed valid arguments:

```{r}
try(calculate_re(daily_input_test,
								 yield = 3000,
								 perennial = FALSE,
								 SoilOrganicC_Percent = 2,
								 ClayContent = 0.05,
								 SandContent = 0.2,
								 
								 tillage_type = "Quadruple tillage",
								 tillage_soil = "Brown"))

try(calculate_re(daily_input_test,
								 yield = 3000,
								 perennial = FALSE,
								 SoilOrganicC_Percent = 2,
								 ClayContent = 0.05,
								 SandContent = 0.2,
								 
								 tillage_type = "Reduced tillage",
								 tillage_soil = "Green"))
```			 
If there is even one NA value in `r_c`, it is estimated using the tillage table.
```{r}
try(calculate_re(daily_input_test,
								 yield = 3000,
								 perennial = FALSE,
								 SoilOrganicC_Percent = 2,
								 ClayContent = 0.05,
								 SandContent = 0.2,
								 
								 r_c = c(NA,rep(0.05,364))))
```


# Running the code over multiple years {#multiyear}

In order to run the script for multiple years, we need a daily input table with a `Year` column.
We can use `Year` as a grouping variable to run `calculate_re` for every year separately.
This creates a vector of r~e~ values the same length as the number of years in the input data.

```{r}
multi_year_input <- 
bind_rows(mutate(daily_input_test, Year = 2020),
					mutate(daily_input_test, Year = 2021),
					mutate(daily_input_test, Year = 2022)
)
multi_year_input
```

```{r}
multi_year_input %>%
	group_by(Year) %>%
	group_split() %>% # Split the table into a list of tables based on the grouping variable
	map(~calculate_re(.,
										 yield = 3000,
										 perennial = FALSE,
										 SoilOrganicC_Percent = 2,
										 ClayContent = 0.05,
										 SandContent = 0.2)) %>% # Run calculate_re for every member of the list
	unlist()
	
```
From the above output, each value in `multi_year_re` corresponds to the yearly value of r~e~.
This vector can then be used as an input parameter in the ICBM R script, which takes a vector of r~e~ values of the same length as the simulation duration.

# References
